Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-23 exp1/iswc2013.tex
Review 2 suggestions
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@613 43ef1190-90fd-4dd7-a769-d0121f02b80e

First, note that the concept of parameters as considered above is
known as \emph{factors} in DoE. For each factor, one may make a
measurement of a \emph{response variable} at certain
\emph{levels}. For example, one may measure the response time of a
SPARQL endpoint by measuring it when it contains 1 or 2 million
triples. Then, the response time is the response variable, the factor
is the number of triples and there are two levels, 1 or 2 million
triples. We have chosen in this paper to only deal with the relatively
straightforward aspects of the DoE formalism, and therefore constrain
the experiment strictly to 2 levels.

@@

A literature survey has not revealed any direct prior art, however, the
general approach has been well established, not only in statistics. A
relatively well cited textbook is \cite{citeulike:5190414} but we have
not found the parts discussed in the present paper to be widely
adopted. Recently, a comprehensive text on experimental methods has
been published in
\cite{Springer-2010-Experimental-Methods-for-the-Analysis-of-Optimization-Algorithms}
demonstrating the broad applicability of the methodology we employ.
We have chosen SPARQL Endpoint evaluation as an example in this study
since SPARQL Federation is our main interest, and to turn to
statistical standard texts \cite{wu2009experiments} for this study.

>>>>

First, note that the concept of parameters as considered above is
known as \emph{factors} in DoE. For each factor, one may make a
measurement of a \emph{response variable} at certain
\emph{levels}. For example, one may measure the response time of a
SPARQL endpoint by measuring it when it contains 1 or 2 million
triples. Levels are not constrained to be continous variables, they
can be discrete, or even two different instances of a class. The
ordering is not essential, but it may contribute to how easy it is to
understand the results, as we shall see. Then, the response time is
the response variable, the factor is the number of triples and there
are two levels, 1 or 2 million triples. We have chosen in this paper
to only deal with the relatively straightforward aspects of the DoE
formalism, and therefore constrain the experiment strictly to 2
levels.

@@

A literature survey has not revealed any direct prior art, neither in
the Semantic Web field nor more generally in database research. However, the
general approach has been well established, not only in statistics. A
relatively well cited textbook is \cite{citeulike:5190414} but we have
not found the parts discussed in the present paper to be widely
adopted. Recently, a comprehensive text on experimental methods has
been published in
\cite{Springer-2010-Experimental-Methods-for-the-Analysis-of-Optimization-Algorithms}
demonstrating the broad applicability of the methodology we employ.
We have chosen SPARQL Endpoint evaluation as an example in this study
since SPARQL Federation is our main interest, and to turn to
statistical standard texts \cite{wu2009experiments} for this study.
